%\documentclass[egilmezThesis.tex]{subfiles} 
%\begin{document}

\chapter*{\centering \LARGE{Acknowledgements}}
\label{chap:Acknowledgements}


\small{This work would not have been possible without the contribution of several people.}

\small{Firstly I would like to thank my supervisors Susana Mu\~noz-Hern\'andez and Victor Pablos Ceruelo for their extreme support, understanding and patience. I really appreciate the fact that they have allowed me to work on a subject, which I sincerely enjoyed and felt passionate about it through the whole process. }

\small{I would like to extend a special thanks to Victor since without his unparalleled input, this work would never be even remotely as good as it's final state. Sometimes in life lady luck smiles at you, and you meet people who are willing to help you with everything they got, regardless of the fact that they do not owe you a penny. Victor is one such unique person and I'm so glad for having the privilege of working with him.}

\small{I'm highly grateful to a list of people starting with prof. H\"{o}lldobler, Dr.Wernhard, Sylvia Wunsch, Paloma Vivas, Julia Koppenhagen and many others for organizing such an amazing program in EMCL. I have spent two wonderful years in two beautiful countries, which I will cherish for the rest of my life.}

\small{It was such a pleasure spending two years with my \textit{compadres}, Alex and Alvaro. Because of you guys I never had a dull moment during our studies.}

\small{I want to thank my dear friends Deniz, Hakan, Serdar and the \textit{projedeyim} gang for making my casual breaks from the thesis study sessions something I eagerly look forward to.}

\small{I owe a huge debt to my high-school teacher Erol Baksi for showing me the path \textit{from zero to hero}. I consider him as a great source of inspiration in my life and can only dream of being able to follow his footsteps one day. }

\small{Last but not least, I must state that I feel immensely indebted to my beloved family, but especially to my mother and my brother Selim for always being there when I needed them, for never having stopped believing in me and for showing me that for every dark night, there's a brighter day.}

%\end{document}