\section{Structured Based Measurement}
\label{sec:SBM}
% The problems of first view, what second view could fix it
By Interpretation Based Measurement in section \ref{sec:IBM}, similarity between predicates is obtained, and used to answer queries in a fuzzy way.  For example, as known, \textit{Beautiful} and \textit{Popular} have a high similarity value, querying \textit{popular(X)}, not only returns results from \textit{popular(X)}, but also from \textit{Beautiful(X)} . In this way, fuzzy query-answering gives more similar choices for users. However, concepts in real world, such as ``geek''  and ``unsociable'' share a large amount of people in each group, by Interpretation Based Measurement, they would be connected with higher similarity value, but semantically, they don't refer to similar concepts according to their linguistics definitions.  In this case, asking for ``geek'', the ``computer scientist''  returns, so does ``unsociable", asking for ``unsociable'', ``shy people'' returns, so does ``geek". Obviously, not only those real similar concepts are gained, but some untruthful results are obtained at the same time, since Interpretation Based Measurement builds strong similarity between ``unsociable'' and ``geek''. In order to avoid this disadvantage, we propose Structured Based Measurement, in which similarity between predicates is defined from their embedded meanings. In RFuzzy Framework, the embedded meaning of predicate is presented by its structure.

In Structured Based Measurement, we build a tree for each predicate, so comparing two predicates is indeed comparing two trees. Section \ref{sec:PredicateTree} introduces the approach to build a tree for predicate. In section \ref{sec:Algorithm}, the algorithm to obtain similarity between two trees is described in detail and its complexity is presented in section \ref{sec:Complexity}. 
\input{Content/Similarity/Content/Structure_Base/Content/predicateTree}
\input{Content/Similarity/Content/Structure_Base/Content/algorithm}
\input{Content/Similarity/Content/Structure_Base/Content/complexity}
