%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%                                                  %%%%%%%%%%%% 
%%%%  SOLUCIONES A LOS PREDICADOS DE PROOF_CNEGF.PL   %%%%%%%%%%%%
%%%%                                                  %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%BASICOS%%%%%%%%%%%%%%%%%%%%%%%%
?- basico1.

yes
?- no_basico1.

no
?- basico2.

no
?- no_basico2.

yes
?- basico3(X).

X = 1 ? ;

no
?- no_basico3(X).

attach_attribute(X,formula(X,[[X/1]])) ? ;

no
?- basico4(X).

attach_attribute(X,formula(X,[[X/_A]])),
attach_attribute(_A,formula(_A,[[X/_A]])) ? ;

no
?- no_basico4(X).

X = fA(_) ? ;

no
?- basico5(X,Y).

Y = X ? ;

no
?- no_basico5(X,Y).

attach_attribute(X,formula(X,[[X/Y]])),
attach_attribute(Y,formula(Y,[[X/Y]])) ? ;

no
?- basico6(X,Y).

attach_attribute(X,formula(X,[[X/Y]])),
attach_attribute(Y,formula(Y,[[X/Y]])) ? ;

no
?- no_basico6(X,Y).

Y = X ? ;

Y = X ? ;

no
?- basico7(X,Y,Z).

Y = X ? ;

Z = X ? ;

no
?- no_basico7(X,Y,Z).

attach_attribute(Z,formula(Z,[[X/Z]])),
attach_attribute(Y,formula(Y,[[X/Y]])),
attach_attribute(X,formula(X,[[X/Y,X/Z]])) ? ;

no
?- basico8(X,Y,Z).

Y = X ? ;

attach_attribute(X,formula(X,[[X/Z]])),
attach_attribute(Z,formula(Z,[[X/Z]])) ? ;

no
?- no_basico8(X,Y,Z).

Z = X,
attach_attribute(X,formula(X,[[X/Y]])),
attach_attribute(Y,formula(Y,[[X/Y]])) ? ;

Z = X,
attach_attribute(X,formula(X,[[X/Y]])),
attach_attribute(Y,formula(Y,[[X/Y]])) ? ;

no

%%%%%%%%%%%%%%%%%%%%%%%%LISTAS%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
?- lista1(X).

X = [1,2,3] ? ;

no
?- no_lista1(X).

attach_attribute(X,formula(X,[[X/[1,2,3]]])) ? ;

no
?- lista2(X,Y).

X = [1,Y,3] ? ;

no
?- no_lista2(X,Y).

attach_attribute(X,formula(X,[[X/[1,Y,3]]])),
attach_attribute(Y,formula(Y,[[X/[1,Y,3]]])) ? ;

no
?- lista3(X,Y,Z).

X = [1,Y,[3,Z,7]] ? ;

no
?- no_lista3(X,Y,Z).

attach_attribute(X,formula(X,[[X/[1,Y,[3,Z,7]]]])),
attach_attribute(Y,formula(Y,[[X/[1,Y,[3,Z,7]]]])),
attach_attribute(Z,formula(Z,[[X/[1,Y,[3,Z,7]]]])) ? ;

no

%%%%%%%%%%%%%%%%%%%%%%%CUNATIFICACION UNIVERSAL%%%%%%%%%%%%%%%%
?- universal1(X).

X = s(_) ? ;

no
?- no_universal1(X).

attach_attribute(X,formula(X,[[X/s(fA(_A))]])),
attach_attribute(_A,formula(_A,[[X/s(fA(_A))]])) ? ;

no
?- universal2(X).

X = s(fA(_)) ? ;

no
?- no_universal2(X).

attach_attribute(X,formula(X,[[X/s(_A)]])),
attach_attribute(_A,formula(_A,[[X/s(_A)]])) ? ;

no
?- universal3(X).

attach_attribute(X,formula(X,[[X/s(_A)]])),
attach_attribute(_A,formula(_A,[[X/s(_A)]])) ? ;

no
?- no_universal3(X).

X = s(fA(_)) ? ;

no
?- universal4(X).

attach_attribute(X,formula(X,[[X/s(fA(_A))]])),
attach_attribute(_A,formula(_A,[[X/s(fA(_A))]])) ? ;

no
?- no_universal4(X).

X = s(_) ? ;

no
?- universal5(X,Y,Z).

X = p(1,Z,_),
attach_attribute(Y,formula(Y,[[Y/q(Z,s(_A),4)]])),
attach_attribute(Z,formula(Z,[[Y/q(Z,s(_A),4)]])),
attach_attribute(_A,formula(_A,[[Y/q(Z,s(_A),4)]])) ? ;

no
?- no_universal5(X,Y,Z).

attach_attribute(X,formula(X,[[X/p(1,Z,fA(_A))]])),
attach_attribute(Z,formula(Z,[[X/p(1,Z,fA(_A))]])),
attach_attribute(_A,formula(_A,[[X/p(1,Z,fA(_A))]])) ? ;

Y = q(Z,s(fA(_)),4) ? ;

Y = q(Z,s(fA(_)),4) ? ;

no

%%%%%%%%%%%%%%%%%%%%%%%%MEZCLAS%%%%%%%%%%%%%%%%%%%%%%%%%%%%
?- mezcla1(X,Y).

X = [1,_,p(2,s(_),[2,_,s(fA(_))])],
Y = 2 ? ;

no
?- no_mezcla1(X,Y).

attach_attribute(X,formula(X,[[X/[1,fA(_A),p(2,s(fA(_D)),[2,fA(_B),s(_C)])]]])),
attach_attribute(_C,formula(_C,[[X/[1,fA(_A),p(2,s(fA(_D)),[2,fA(_B),s(_C)])]]])),
attach_attribute(_A,formula(_A,[[X/[1,fA(_A),p(2,s(fA(_D)),[2,fA(_B),s(_C)])]]])),
attach_attribute(_D,formula(_D,[[X/[1,fA(_A),p(2,s(fA(_D)),[2,fA(_B),s(_C)])]]])),
attach_attribute(_B,formula(_B,[[X/[1,fA(_A),p(2,s(fA(_D)),[2,fA(_B),s(_C)])]]])) ? ;

attach_attribute(Y,formula(Y,[[Y/2]])) ? ;

no
?- mezcla2(X,Y,Z,W,F).

W = [_,0] ? ;

X = p(W,s(Y),F,_) ? ;

F = s(_A),
attach_attribute(Z,formula(Z,[[Z/s(_A)]])),
attach_attribute(_A,formula(_A,[[Z/s(_A)]])) ? ;

no
?- no_mezcla2(X,Y,Z,W,F).

Z = s(fA(_)),
attach_attribute(W,formula(W,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])),
attach_attribute(_A,formula(_A,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])),
attach_attribute(X,formula(X,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])),
attach_attribute(F,formula(F,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])),
attach_attribute(Y,formula(Y,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])),
attach_attribute(_B,formula(_B,[[W/[fA(_A),0],X/p(W,s(Y),F,fA(_B))]])) ? ;

attach_attribute(F,formula(F,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(_A,formula(_A,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(W,formula(W,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(_C,formula(_C,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(X,formula(X,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(Y,formula(Y,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])),
attach_attribute(_B,formula(_B,[[F/s(fA(_A)),X/p(W,s(Y),F,fA(_B)),W/[fA(_C),0]]])) ? ;

no
?- mezcla3(X,Y,Z,W).

X = 1,
Y = 2,
attach_attribute(Z,formula(Z,[[Z/3]])),
attach_attribute(W,formula(W,[[W/4]])) ? ;

no
?- no_mezcla3(X,Y,Z,W).

attach_attribute(X,formula(X,[[X/1]])) ? ;

attach_attribute(Y,formula(Y,[[Y/2]])) ? ;

Z = 3 ? ;

W = 4 ? ;

no
?- mezcla4(X).

attach_attribute(X,formula(X,[[X/3,X/4]])) ? ;

no
?- no_mezcla4(X).

X = 3 ? ;

X = 4 ? ;

no
?- mezcla5(X,Y).

attach_attribute(Y,formula(Y,[[X/Y],[Y/2]])),
attach_attribute(X,formula(X,[[X/3,X/Y],[X/3,Y/2]])) ? ;

no
?- no_mezcla5(X,Y).

X = 3 ? ;

X = 3,
Y = 2 ? ;

X = 3,
Y = 3 ? ;

X = 2,
Y = 2 ? ;

X = 2,
Y = 2 ? ;

no
?- mezcla6(X,Y).

attach_attribute(_A,formula(_A,[[X/p(3,_A)]])),
attach_attribute(Y,formula(Y,[[X/1],[Y/2]])),
attach_attribute(X,formula(X,[[X/1,X/p(3,_A)],[Y/2,X/p(3,_A)]])) ? ;

no
?- no_mezcla6(X,Y).

X = 1,
Y = 2 ? ;

X = p(3,fA(_)),
Y = 2 ? ;

X = p(3,fA(_)) ? ;

X = 1,
Y = 2 ? ;

no
?- mezcla7(X,Y,Z,W).

X = negacion,
Y = 0,
attach_attribute(Z,formula(Z,[[W/s(s(fA(Z)))]])),
attach_attribute(W,formula(W,[[W/s(s(fA(Z)))]])) ? ;

X = negacion,
Y = s(0),
attach_attribute(Z,formula(Z,[[W/s(s(fA(Z)))]])),
attach_attribute(W,formula(W,[[W/s(s(fA(Z)))]])) ? ;

no
?- no_mezcla7(X,Y,Z,W).

attach_attribute(X,formula(X,[[X/negacion]])) ? ;

attach_attribute(X,formula(X,[[X/negacion]])),
attach_attribute(Y,formula(Y,[[Y/s(0)]])) ? ;

W = s(s(Z)),
attach_attribute(X,formula(X,[[X/negacion]])) ? ;

W = s(s(Z)),
attach_attribute(X,formula(X,[[X/negacion]])) ? ;

attach_attribute(Y,formula(Y,[[Y/0]])),
attach_attribute(X,formula(X,[[X/negacion]])) ? ;

attach_attribute(Y,formula(Y,[[Y/s(0),Y/0]])) ? ;

W = s(s(Z)),
attach_attribute(Y,formula(Y,[[Y/0]])) ? ;

W = s(s(Z)),
attach_attribute(Y,formula(Y,[[Y/0]])) ? ;

W = s(s(Z)),
attach_attribute(X,formula(X,[[X/negacion]])) ? ;

W = s(s(Z)),
attach_attribute(Y,formula(Y,[[Y/s(0)]])) ? ;

W = s(s(Z)) ? ;

W = s(s(Z)) ? ;

W = s(s(Z)),
attach_attribute(X,formula(X,[[X/negacion]])) ? ;

W = s(s(Z)),
attach_attribute(Y,formula(Y,[[Y/s(0)]])) ? ;

W = s(s(Z)) ? ;

W = s(s(Z)) ? ;

no
?- mezcla8(X,Y,Z).

attach_attribute(X,formula(X,[[X/Z],[Y/5]])),
attach_attribute(Y,formula(Y,[[X/Z],[Y/5]])),
attach_attribute(Z,formula(Z,[[X/Z,Z/6],[Y/5,Z/6]])) ? ;

no
?- no_mezcla8(X,Y,Z).

Y = 5,
Z = X ? ;

Y = 5,
Z = X ? ;

Y = 5,
Z = X ? ;

X = 6,
Z = 6 ? ;

Y = 5,
Z = 6 ? ;

Z = 6 ? ;

no
?- mezcla9(X,Y,Z,W,F,G).

F = s(fA(_A)),
G = s(_B),
X = [[W,W],2],
Y = [W,W],
Z = 2,
attach_attribute(_B,formula(_B,[[W/[s(fA(_A)),s(_B)]]])),
attach_attribute(_A,formula(_A,[[W/[s(fA(_A)),s(_B)]]])),
attach_attribute(W,formula(W,[[W/[s(fA(_A)),s(_B)]]])) ? ;

no
?- no_mezcla9(X,Y,Z,W,F,G).

attach_attribute(X,formula(X,[[X/[[W,W],2]]])),
attach_attribute(W,formula(W,[[X/[[W,W],2]]])) ? ;

attach_attribute(Y,formula(Y,[[Y/[W,W]]])),
attach_attribute(W,formula(W,[[Y/[W,W]]])) ? ;

attach_attribute(Z,formula(Z,[[Z/2]])) ? ;

W = [s(_),s(fA(_))] ? ;

attach_attribute(F,formula(F,[[F/s(_A)]])),
attach_attribute(_A,formula(_A,[[F/s(_A)]])) ? ;

attach_attribute(G,formula(G,[[G/s(fA(_A))]])),
attach_attribute(_A,formula(_A,[[G/s(fA(_A))]])) ? ;

no

